Differential neural ensemble search with diversity control

### 2023-04-12 paper/main.tex
paper updated

\usepackage[utf8]{inputenc}
\usepackage[english, russian]{babel}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{float}
\usepackage{wrapfig}
\graphicspath{{../figures}}
\usepackage[square,numbers]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{doi}

@@

In order to control diversity we employ the concept of hypernetwork. A hypernetwork is a parametric mapping from $[0, \Lambda]$ to the set of model structural parameters \citep{darts-cc}.
$$
\mbox{\boldmath{$\alpha$}} : [0, \Lambda], \mathbb{R}^u \to \mathbb{R}^s
$$
Where $\mathbb{R}^u$ is a hypernetwork parametric space and $\mathbb{R}^s$ is space of model structural parameters. In this terms \mbox{\boldmath{$\alpha$}} can be redefined using hypernetwork

@@

\begin{gather*}
    \min_{\mbox{\boldmath{$\alpha$}}} \mathbb{E}_{\lambda \sim U(0, \Lambda)} [\mathcal{L}_{val}(\mbox{\boldmath{$w^*$}}, \mbox{\boldmath{$\alpha$}}(\lambda)) - \lambda JS(\mbox{\boldmath{$\alpha^*$}}, \mbox{\boldmath{$\alpha$}}(\lambda))] \\
    s.t. \text{ } \mbox{\boldmath{$w^*$}} = \arg \min_{\mbox{\boldmath{$w$}}} \mathbb{E}_{\lambda \sim U(0, \Lambda)}[\mathcal{L}_{train}(\mbox{\boldmath{$w$}}, \mbox{\boldmath{$\alpha$}}(\lambda))]
\end{gather*}

Following problem resolution, a new architecture is acquired and subsequently included into the ensemble.
It is assured that said architecture is diverse enough from its counterparts and performs an acceptable level of accuracy.

@@

First goal of computational experiment is to investigate dependence of resulting architecture's performance on $\lambda$ and also to estimate suitable $\Lambda$.
We conducted experiments on fashionMNIST dataset for $\lambda = 1, \ldots, 40$ and looked at resulting architecture and accuracy.
Second goal is to investigate effectiveness of ensembling.
We investigated how to choose architectures that on one hand are diverse enough and on other hand perform accaptable accuracy.

@@

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}{.5\textwidth}
	  \centering
	  \includegraphics[width=1\linewidth]{fig2}
	  \caption{Dependence of accuracy on diverse parameter $\lambda$}
	  \label{fig:graph1}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
	  \centering
	  \includegraphics[width=1\linewidth]{fig3}
	  \caption{Dependence of matching edges number on diverse parameter $\lambda$}
	  \label{fig:graph2}
	\end{subfigure}
	\caption{Graphical results of computational experiment}
	\label{fig:graph}
\end{figure}

@@

\begin{table}[h   ]
	\caption{preliminary results}
	\centering
	\begin{tabular}{lll}
		\midrule
		$\lambda$     & accuracy, \% & matched operations \\
		\midrule
        0 (optimum) & 91.24 & - \\ 
        1/2  & 91.24 &  4   \\
		2  & 91.26   &  3   \\
		32 & 89.21   &  2   \\
		64 & 87.09   &  1   \\
		\bottomrule
	\end{tabular}
	\label{tab:prelim}
\end{table}

@@

To investigate the effectiveness of ensembling, we employed the following methodology.
We randomly sampled several architectures corresponding to different diverse parameters $\lambda$ chosen randomly and uniformaly from a range of 0 to $\Lambda$.
The obtained architectures were then included in an ensemble, and the accuracy of the resulting ensemble was measured.
By changing the diversity limit $\Lambda$, we were able to control the diversity of our deep learning models.
The graphical results of our experiment are depicted in \hyperref[fig:graph3]{Fig.3}.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{fig4}
    \caption{\label{fig:graph3} Dependance of effectiveness of deep learning models ensemble on diversity limit $\Lambda$}
\end{figure}

@@

In this paper, we proposed a novel method for sampling ensembles of deep learning models with diversity control.
Our method utilizes a hypernetwork to generate diverse architectures by perturbing a base architecture in terms of Jensen-Shannon divergence.
The diversity of the ensemble is controlled by a penalty term added to the loss function, which encourages the ensemble members to be diverse.
We conducted experiments on the Fashion-MNIST dataset and demonstrated that our method performs compatible results in terms of accuracy and architectural diversity.
Overall, our proposed method shows potential for practical applications in deep learning ensembling. asddasd

>>>>

\usepackage[utf8]{inputenc}
\usepackage[english, russian]{babel}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{float}
\usepackage{wrapfig}
\graphicspath{{../figures}}
\usepackage[square,numbers]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{doi}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amssymb}

\usepackage{doi}

@@

In order to control diversity we employ the concept of hypernetwork. A hypernetwork is a parametric mapping from $[0, \Lambda]$ to the set of model structural parameters \citep{darts-cc}.
$$
\mbox{\boldmath{$\alpha$}} : [0, \Lambda] \times \mathbb{R}^u \to \mathbb{R}^s
$$
Where $\mathbb{R}^u$ is a hypernetwork parametric space and $\mathbb{R}^s$ is space of model structural parameters. In this terms \mbox{\boldmath{$\alpha$}} can be redefined using hypernetwork

@@

\begin{gather*}
        \min_{\mbox{\boldmath{$\alpha$}}} \mathbb{E}_{\lambda \sim U(0, \Lambda)} [\mathcal{L}_{val}(w^*, \mbox{\boldmath{$\alpha$}}(\lambda)) - \lambda JS(\mbox{\boldmath{$\alpha^*$}}, \mbox{\boldmath{$\alpha$}}(\lambda))] \\
    s.t. \text{ } \mbox{\boldmath{$w^*$}} = \arg \min_{\mbox{\boldmath{$w$}}} \mathbb{E}_{\lambda \sim U(0, \Lambda)}[\mathcal{L}_{train}(\mbox{\boldmath{$w$}}, \mbox{\boldmath{$\alpha$}}(\lambda))]
\end{gather*}

Following problem resolution, a new architecture is acquired and subsequently included into the ensemble. It is assured that said architecture is diverse enough from its counterparts and performs an acceptable level of accuracy. Formal algorithm is written down as \hyperref[alg:1]{Algorithm 1}.
\begin{figure}[h]  
  \begin{minipage}{.6\linewidth}
    \begin{algorithm}[H]
    \caption{}\label{alg:1}
        \begin{algorithmic}
                \State Initialize: $N \in \mathbb{N}$, $\mathcal{S} = \varnothing$, hypernetwork
            \State $S \gets \{ \mbox{\boldmath{$\alpha^*$}} \}$ \Comment{Result os NAS}
            \For{$i = 1, \ldots, N$}
                \State Sample $\lambda \sim U(0, \Lambda)$
                \State $\mathcal{S} \gets \mathcal{S} \cup \{ \alpha(\lambda, \mbox{\boldmath{$\alpha^*$}}) \}$ \Comment{$\alpha$ gaind from hypernetwork}
            \EndFor
            \State \textbf{Return:} S as a resulting ensemble
        \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\end{figure}

@@

In this study, we conducted a series of experiments on the fashionMNIST dataset \citep{xiao2017fashion} using a convolutional neural network (CNN) model.
Specifically, we utilized a one-layer architecture with six nodes and seven possible operations for each edge, as outlined in the DARTS methodology \cite{darts}.
Our optimizer was stochastic gradient descent (SGD), and we employed a cosine annealing learning rate scheduler with a minimum learning rate of 0.001. The loss criterion utilized in this study was CrossEntropyLoss.

First goal of computational experiment is to investigate dependence of resulting architecture's performance and their diversity on $\lambda$.
Second goal is to investigate effectiveness of ensembling.
Main question is how to choose architectures that on one hand are diverse enough and on other hand perform acceptable accuracy.

@@

We employed the Differentiable Architecture Search (DARTS) algorithm to sample a base architecture with $\lambda = 0$. Subsequently, we sampled architectures for values of $\lambda$ ranging from 1 to 40, recording the resulting architectural structures and their corresponding classification performance. Specifically, we sought to examine the impact of varying levels of architectural diversity on model accuracy, with the aim of identifying optimal architectures for the given task.

@@

\begin{table}[h   ]
	\centering
	\begin{tabular}{lll}
		\midrule
		$\lambda$     & accuracy, \% & matched operations \\
		\midrule
        0 (optimum) & 92.10 & 18 \\ 
        5  & 89.66 &  2   \\
		10  & 90.59   &  2   \\
		15 & 90.09   &  1   \\
		20 & 88.52   &  1   \\
      25 & 88.54   &  1   \\
      30 & 87.42   &  2   \\
      35 & 86.41   &  1   \\
      40 & 88.67   &  2  \\
      \hline
    \vspace{0.1cm}      
	\end{tabular}
    \caption{Table to test captions and labels.}
	\label{tab:prelim}
\end{table}

\begin{figure}[h]
	\centering
	\begin{subfigure}{.46\textwidth}
        \centering
	    \includegraphics[width=1\linewidth]{fig2}
	  \caption{Dependence of performed accuracy on diversity parameter $\lambda$}
	  \label{fig:graph1}
	\end{subfigure}%
    \hspace{1cm}
	\begin{subfigure}{.46\textwidth}
	  \centering
	  \includegraphics[width=1\linewidth]{fig3}
	  \caption{Dependence of matching edges number on diversity parameter $\lambda$}
	  \label{fig:graph2}
	\end{subfigure}
	\caption{Graphical results of computational experiment}
	\label{fig:graph}
\end{figure}

@@

@@

To investigate the effectiveness of ensembling, we employed the following methodology.
We randomly sampled three architectures corresponding to different diverse parameters $\lambda$ chosen randomly and uniformly from a range of 0 to $\Lambda$.
The obtained architectures were then included in an ensemble, and the accuracy of the resulting ensemble was measured.
By changing the diversity limit $\Lambda$, we were able to control the diversity of our deep learning models.
The graphical results of our experiment are depicted in \hyperref[fig:graph3]{Fig.3}.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.55\textwidth]{fig4}
    \caption{\label{fig:graph3} Dependence of effectiveness of deep learning models ensemble on diversity limit $\Lambda$}
\end{figure}

@@

In this paper, we proposed a novel method for sampling ensembles of deep learning models with diversity control. Our method utilizes a hypernetwork to generate diverse architectures by perturbing a base architecture in terms of Jensen-Shannon divergence. The diversity of the ensemble is controlled by a penalty term added to the loss function, which encourages the ensemble members to be diverse. We conducted experiments on the Fashion-MNIST dataset and demonstrated that our method performs compatible results in terms of accuracy and architectural diversity. Overall, our proposed method shows potential for practical applications in deep learning ensembling.
